Best practices for forecasting changes in biodiversity: an example using breeding birds

### 2017-08-30 manuscript/draft.Rmd
Minor edits to the box
1. Change title to "Best practices" since that's how they are described in the
   paper
2. Add some references
3. Minor editorial work

## Box 1: Ten simple rules for making and evaluating ecological forecasts

@@

Typically ecological forecasts use one modeling approach or a small number of
related approaches. By fitting and evaluating multiple modeling approaches we
can learn more rapidly about the best approaches for making predictions for a
given ecological quantity. This includes comparing process-based [e.g.
@kearney2009] and data-driven models, as well as comparing the accuracy of
forecasts to simple baselines to determine if the modeled forecasts are more
accurate than the naive assumption that the world is static [@ye2015].

@@

Forecasts describe how systems are expected to change through time. While some
areas of ecological forecasting focus primarily on time-series data, others
primarily focus on using spatial models and space-for-time substitutions. Using
ecological and environmental time-series data allows the consideration of actual
dynamics from both a process and error structure perspective [@tredennick2016].

@@

Understanding uncertainty in a forecast is just as important as understanding
the average or expected outcome. Failing to account for uncertainty can result
in overconfidence in highly uncertain outcomes leading to poor decision making
and erosion of confidence in ecological forecasts. Models should explicitly
include sources of uncertainty and propagate them through the forecast where
possible [@clark2001; @dietze2017]. Evaluations of forecasts should assess the
accuracy of models' estimated uncertainties as well as their point estimates
[@dietze2017].

@@

In general, the accuracy of forecasts decreases with the length of time into the
future being forecast [@petchey2015]. This decay in accuracy (and the potential
for different rates of decay to result in different relative model performance
at different lead times) should be considered when evaluating forecasts
and comparing models.

@@

Ecological observations are influenced by both the underlying biological
processes (e.g. resource limitation) and how the system is sampled. When
possible, forecasts should model the factors influencing the observation of the
data. [@yu2010; @hutchinson2011; @schurr2012]

@@

Evalutating a model's predictive performance across time is critical.
Hindcasting uses a temporal out-of-sample validation approach to mimic how well
a model would have performed had it been run in the past. For example, using
occurance data from the early 20th century to model distributions which are
validated with late 20th century occurances. Dense time series, such as yearly
observations, are desirable to also evalulate the forecast horizon (see #6), but
this is not a strict requirement.

@@

Forecast values and/or models should be archived so that they can be assessed
after new data is generated [@mcgill2012; @silver2012]. Enough information
should be provided in the archive to allow ecologists to make an unambiguous
assessment of each forecast's performance.

@@

In cases where long-term predictions are the primary goal, short-term
predictions should also be made to accommodate the time-scales of planning and
management decisions and to allow the accuracy of the forecasts to be quickly
evaluated [@tredennick2016].

>>>>

## Box 1: Best practices for making and evaluating ecological forecasts

@@

Typically ecological forecasts use one modeling approach or a small number of
related approaches. By fitting and evaluating multiple modeling approaches we
can learn more rapidly about the best approaches for making predictions for a
given ecological quantity [@clark2001; @ward2014]. This includes comparing
process-based [e.g., @kearney2009] and data-driven models [e.g., @ward2014], as
well as comparing the accuracy of forecasts to simple baselines to determine if
the modeled forecasts are more accurate than the naive assumption that the world
is static [@jolliffe2003; @ye2015].

@@

Forecasts describe how systems are expected to change through time. While some
areas of ecological forecasting focus primarily on time-series data [@ward2014],
others primarily focus on using spatial models and space-for-time substitutions
[ref?]. Using ecological and environmental time-series data allows the
consideration of actual dynamics from both a process and error structure
perspective [@tredennick2016].

@@

Understanding uncertainty in a forecast is just as important as understanding
the average or expected outcome. Failing to account for uncertainty can result
in overconfidence in uncertain outcomes leading to poor decision making and
erosion of confidence in ecological forecasts [@clark2001]. Models should
explicitly include sources of uncertainty and propagate them through the
forecast where possible [@clark2001; @dietze2017]. Evaluations of forecasts
should assess the accuracy of models' estimated uncertainties as well as their
point estimates [@dietze2017].

@@

In general, the accuracy of forecasts decreases with the length of time into the
future being forecast [@petchey2015]. This decay in accuracy should be
considered when evaluating forecasts. In addition to simple decreases in
forecast accuracy the potential for different rates of decay to result in
different relative model performance at different lead times should be
considered.

@@

Ecological observations are influenced by both the underlying biological
processes (e.g. resource limitation) and how the system is sampled. When
possible, forecasts should model the factors influencing the observation of the
data [@yu2010; @hutchinson2011; @schurr2012].

@@

Evalutating a model's predictive performance across time is critical for
understanding if it is useful for forecasting the future. Hindcasting uses a
temporal out-of-sample validation approach to mimic how well a model would have
performed had it been run in the past. For example, using occurance data from
the early 20th century to model distributions which are validated with late 20th
century occurances. Dense time series, such as yearly observations, are
desirable to also evalulate the forecast horizon (see #6), but this is not a
strict requirement.

@@

Forecast values and/or models should be archived so that they can be assessed
after new data is generated [@mcgill2012; @silver2012; @dietze2016]. Enough
information should be provided in the archive to allow unambiguous assessment of
each forecast's performance [@tetlock2016].

@@

Even in cases where long-term predictions are the primary goal, short-term
predictions should also be made to accommodate the time-scales of planning and
management decisions and to allow the accuracy of the forecasts to be quickly
evaluated [@dietze2016; @tredennick2016].
